---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r global_options, include=FALSE} 

knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.path='Figs/',
                      echo=TRUE, warning=FALSE, message=FALSE)
options(digits=2)
options(scipen=999)
```

### Loading and preprocessing the data
```{r Loading_Data}
setwd("C:/Users/Taaja/Desktop/R Reproducible Research/RepData_PeerAssessment1")
act <- read.csv("activity.csv")
```

Here is a histogram of the total number of steps taken each day.

```{r Histo}
sumperday <- aggregate(steps ~ date, data = act, sum)
hist(sumperday$steps, xlab="Steps", main = "Histogram of the number of steps")
```

###Mean and median steps
On average, this person had `r mean(sumperday$steps, na.rm = TRUE)` steps per day (median: `r median(sumperday$steps, na.rm = TRUE)` steps).

If you want to see the code used to calculate those numbers, this was the code I used in-line:
```{r means}
#mean(sumperday$steps, na.rm = TRUE)
#median(sumperday$steps, na.rm = TRUE)
```

### What is the average daily activity pattern?
```{r plot_activity}
act.NA <- act[complete.cases(act),]
av.int <- aggregate(steps ~ interval, data = act.NA, mean, na.rm=TRUE)
plot(av.int$interval, av.int$steps, type="l", xlab="Interval", ylab="Average steps")
```

Which 5-minute interval, on average across all the days in the dataset, 
contains the maximum number of steps?

The `r av.int$interval[max(av.int$steps)]` interval had the maximum number of steps. 

If you want to see the code used to calculate those numbers, this was the code I used in-line::
```{r int.max}
#av.int$interval[max(av.int$steps)]
```

### Imputing missing values
In total, there were `r sum(is.na(act))` missing values in the dataset. I replaced the missing values with a glm model of interval and daynumber. 

```{r fillin}
#in-line above:
#r sum(is.na(act)) Find the number of missing values

library(lubridate)
act.NA$daynum <- yday(as.Date(act.NA$date))
model <- glm(steps ~ interval + daynum, data=act.NA)

missing <- act[is.na(act),]
missing$steps <- NULL
missing$daynum <- yday(as.Date(missing$date))

preds <- predict(model, missing[,2:3], se = TRUE)

missing$steps <- preds[["fit"]]

act.preds<- rbind(missing, act.NA)

sumperday.preds <- aggregate(steps ~ date, data = act.preds, sum)
```

Here is a histogram with the replaced missing values.
```{r pred_histo}
hist(sumperday.preds$steps, xlab="Steps", main = "Histogram of the number of steps")
```

###Mean and median number of steps (with replacement values)
On average, this person had `r mean(sumperday.preds$steps, na.rm = TRUE)` steps per day (median: `r median(sumperday.preds$steps, na.rm = TRUE)` steps).

```{r means_replace}
#mean(sumperday.preds$steps, na.rm = TRUE)
#median(sumperday.preds$steps, na.rm = TRUE)
```


### Are there differences in activity patterns between weekdays and weekends?
Yes! Check out this plot.

```{r fillinresults}
act.preds$date <- as.Date(act.preds$date)
act.preds$weekday <- weekdays(act.preds$date)
act.preds$weekday.or.weekend <- ifelse(act.preds$weekday %in% c("Sunday", "Saturday"), "weekend","weekday")

av.int.pred <- aggregate(steps ~ interval + weekday.or.weekend, data = act.preds, mean, na.rm=TRUE)

library(lattice)
xyplot(steps~interval|weekday.or.weekend, data=av.int.pred, type ="l",
       main="Average steps for weekends and weekdays", 
       ylab="Average steps", xlab="Time interval")
```
